### 2014-05-31 ACL2014/abstract.tex
Added updated latex file

\begin{abstract}
Many Web, Mobile, and NLP applications 
rely on finding nearest neighbors 
for a given query from a commercial large-scale query logs . 
Due to the large scale volume of queries, 
computing nearest neighbors from all queries using brute force linear 
search is a computationally time intensive task. 
To challenge the aforementioned problem, we exploit existing 
\lshf (a.k.a, \lsh) methods and propose 
their novel variants in distributed setting (e.g. Hadoop). 
Our experimental results show that couple of our 
new variants of \lsh get signiﬁcantly better recall than vanilla \lsh by 
using the same number of hash tables. 
\end{abstract}

>>>>

\begin{abstract}
Many Web, Mobile, and NLP applications 
rely on finding nearest neighbors 
for a given query from a commercial large-scale query logs . 
Due to the large scale volume of user queries, 
computing nearest neighbors from all queries using brute force linear 
search is a computationally time intensive task. 
To challenge the aforementioned problem, we exploit existing 
\lshf (a.k.a, \lsh) methods and propose 
their novel variants in distributed setting (e.g. Hadoop). 
Our experimental results show that two of our 
new variants of \lsh get significantly better recall than vanilla \lsh by 
using the same number of hash tables. 
\end{abstract}

### 2014-05-31 ACL2014/experiments.tex
Added updated latex file

\subsection{Evaluating Vanilla LSH} 
\label{subsec:eval:vanillaLSH}
In the first experiment, we vary $L=\{1,10,28,55\}$ with fixed $K=16$. 
$L$ denotes the number of hash tables and $K$ denotes the length of the index of the buckets in the table. 
If we increase $K$ (increasing the precision to reduce false positives), we also need to 
increase $L$ to get good recall (increasing the recall to reduce false negatives). 
Table \ref{tab:varyL} shows that increasing $L$ leads to better recall, 
however at the expense of more comparisons. In addition, having large $L$ 
means generating large number of random projection bits and hash tables that is 
both time and memory intensive. Hence, we fix $L=10$, 
which leads to reasonable recall with fewer comparisons.

In the second experiment, we vary $K=\{4,8,16\}$ with fixed $L=10$. 
As expected, Table \ref{tab:varyK} shows that increasing $K$ reduces the number of 
comparisons and worse recall. This is intuitive as  
the larger value of $K$ leads to larger 
gap between probabilites of collision between close queries 
and far queries (see Section \ref{sec:vlsh}). 
Hence, we fix $K=16$ to have fewer number of comparisons.

In the third experiment, we fix $L=10$ and $K=16$ and increase the size of training data. 
Table \ref{tab:varyData} demonstrates that as we increase the training data size, 
the number of comparisons also increase. This result indicates that $K$   
needs to be tuned with respect to a specific dataset, 
as large $K$ will reduce the probability of dis-similar 
queries falling within the same bucket. 
$K$ and $L$ can be tuned by 
randomly sampling small set of queries.\footnote{In this paper, we randomly select 2000 queries to tune parameter $K$.}

In the fourth experiment, Table \ref{tab:bestLSH} shows the best $K$ and $L$ parameter settings on different 
sized datasets.\footnote{Recall on \dataC cannot be computed, as it was computationally intensive to find exact similar neighbors.} 
On our biggest dataset of 600 million queries, we set $K=24$, $L=10$. These parameter settings require only 
$464$ comparisons to find approximate nearest neighbors compared to exact cosine similarity that involves 
brute force search over all 600 million queries in the dataset.

\begin{table}
\centering
\begin{tabular}{|c|r|c|}
\hline
$L$ & Comparisons & Recall \\
\hline
1 & 7  & .28 \\
 \rowcolor[gray]{0.9} 10  &  57 & .63 \\
28 &  152 & .77 \\
55 &  297 & .89 \\
\hline 
 \end{tabular}
\caption{\footnotesize{Varying $L$ with fixed $K=16$ on \aol with $\tau=0.7$.}}
\label{tab:varyL}
\end{table}

\begin{table}
\centering
\begin{tabular}{|c|r|c|}
\hline
$K$ & Comparisons & Recall \\
\hline
4  &  112,347 & .98 \\
8 &  11,008 & .90 \\
\rowcolor[gray]{0.9}16 &  57 & .63 \\
\hline 
 \end{tabular}
\caption{\footnotesize{Varying $K$ with fixed $L=10$ on \aol with $\tau=0.7$.}}
\label{tab:varyK}
\end{table}

\begin{table}
\centering
\begin{tabular}{|c|r|c|}
\hline
Data & Comparisons & Recall \\
\hline
\aol & 57  & .63 \\
\dataA  &  1,052 & .67 \\
\dataB    & 10,515 & .64 \\
\dataC    & - & - \\
\hline 
 \end{tabular}
\caption{\footnotesize{Fixed $K=16$  and $L=10$ on different sized datasets with $\tau=0.7$.}}
\label{tab:varyData}
\end{table}

>>>>

\subsection{Evaluating Vanilla LSH} 
\label{subsec:eval:vanillaLSH}

\begin{table}
\centering
\begin{tabular}{|c|rc|rc|}
\hline
\multirow{2}{*}{$\tau$} & \multicolumn{2}{c}{\aol} & \multicolumn{2}{c}{\dataA} \\
 & Comparisons & Recall & Comparisons & Recall  \\
\hline
0.7 & \multirow{3}{*} {57}  & .63  & \multirow{3}{*} {1052} & .67 \\
0.8 &   & .84  &  & .81 \\
0.9 &   & .98  &  & .96 \\
\hline 
 \end{tabular}
\caption{\footnotesize{Varying $\tau$ with fixed $K=16$ and $L=10$ on \aol and \dataA.}}
\label{tab:varyTau}
\end{table}

In the first experiment, we vary the similarity threshold parameter $\tau={0.7,0.8,0.9}$ 
with fixed $K=16$ and $L=10$ on \aol and \dataA datasets. 
Table \ref{tab:varyTau} shows that as expected finding near-duplicates ($0.9$) 
is easier than nearest neighbors ($0.7$). For, rest of this paper, 
we fix $\tau=0.7$ as we are interested in both near-duplicates and nearest-neighbors.

In the second experiment, we vary $L=\{1,10,28,55\}$ with fixed $K=16$ on \aol and \dataA datasets. 
$L$ denotes the number of hash tables and $K$ denotes the length of the index of the buckets in the table. 
If we increase $K$ (increasing the precision to reduce false positives), we also need to 
increase $L$ to get good recall (increasing the recall to reduce false negatives). 
Table \ref{tab:varyL} shows that increasing $L$ leads to better recall, 
however at the expense of more comparisons on both the datasets. 
In addition, having large $L$ 
means generating large number of random projection bits and hash tables that is 
both time and memory intensive. Hence, we fix $L=10$, 
which leads to reasonable recall with fewer comparisons.

\begin{table}
\centering
\begin{tabular}{|c|rc|rc|}
\hline
\multirow{2}{*}{L} & \multicolumn{2}{c}{\aol} & \multicolumn{2}{c}{\dataA} \\
 & Comparisons & Recall & Comparisons & Recall  \\
\hline
1 & 7  & .28 & 106 & .36 \\
 \rowcolor[gray]{0.9} 10  &  57 & .63 & 1052 & .67 \\
28 &  152 & .77 & 2908 & .78 \\
55 &  297 & .89 & 5648 & .84 \\
\hline 
 \end{tabular}
\caption{\footnotesize{Varying $L$ with fixed $K=16$ on \aol and \dataA with $\tau=0.7$.}}
\label{tab:varyL}
\end{table}

In the third experiment, we vary $K=\{4,8,16\}$ with fixed $L=10$ on \aol and \dataA datasets. 
As expected, Table \ref{tab:varyK} shows that increasing $K$ reduces the number of 
comparisons and worse recall on both the datasets. This is intuitive as  
the larger value of $K$ leads to larger 
gap between probabilites of collision between close queries 
and far queries (see Section \ref{sec:vlsh}). 
Hence, we fix $K=16$ to have fewer number of comparisons.

\begin{table}
\centering
%\begin{tabular}{|c|r|c|}
\begin{tabular}{|c|rc|rc|}
\hline
%$K$ & Comparisons & Recall \\
\multirow{2}{*}{K} & \multicolumn{2}{c}{\aol} & \multicolumn{2}{c}{\dataA} \\
 & Comparisons & Recall & Comparisons & Recall  \\
\hline
4  &  112,347 & .98 & 2,29,2670 & .96 \\
8 &  11,008 & .90 & 221,132 & .88 \\
\rowcolor[gray]{0.9}16 &  57 & .63 & 1,052 & .67 \\
\hline 
 \end{tabular}
\caption{\footnotesize{Varying $K$ with fixed $L=10$ on \aol with $\tau=0.7$.}}
\label{tab:varyK}
\end{table}

In the fourth experiment, we fix $L=10$ and $K=16$ and increase the size of training data. 
Table \ref{tab:varyData} demonstrates that as we increase the training data size, 
the number of comparisons also increase. This result indicates that $K$   
needs to be tuned with respect to a specific dataset, 
as large $K$ will reduce the probability of dis-similar 
queries falling within the same bucket. 
$K$ and $L$ can be tuned by 
randomly sampling small set of queries.\footnote{In this paper, we randomly select 2000 queries to tune parameter $K$.}

In the fifth experiment, Table \ref{tab:bestLSH} shows the best $K$ and $L$ parameter settings on different 
sized datasets.\footnote{Recall on \dataC cannot be computed, as it was computationally intensive to find exact similar neighbors.} 
On our biggest dataset of 600 million queries, we set $K=24$, $L=10$. These parameter settings require only 
$464$ comparisons to find approximate nearest neighbors compared to exact cosine similarity that involves 
brute force search over all 600 million queries in the dataset.

\begin{table}
\centering
\begin{tabular}{|c|r|c|}
\hline
Data & Comparisons & Recall \\
\hline
\aol & 57  & .63 \\
\dataA  &  1,052 & .67 \\
\dataB    & 10,515 & .64 \\
\dataC    & 105,126 & - \\
\hline 
 \end{tabular}
\caption{\footnotesize{Fixed $K=16$  and $L=10$ on different sized datasets with $\tau=0.7$.}}
\label{tab:varyData}
\end{table}

### 2014-05-31 ACL2014/intro.tex
Updated latex files
Added updated latex file

 \section{Introduction}
Millions of users visit commercial search engines and ``query'' 
their interests. To provide users with high quality of services, 
search engines such as Bing, Google, and Yahoo require
intelligent analysis to realize users' implicit intents, in particular taking
advantage of very large scaled query logs.  One of the key
interesting tasks involved in learning users' implicit intents often involves 
computing nearest neighbors (queries) for a user given query from all queries. 
Computing nearest neighbors is useful for 
many search-related problems on the Web and the Mobile such as 
finding related queries \cite{Jones06WWW,Jain11SIGIR,Song12WSDM}, 
finding near-duplicate queries \cite{}, spelling correction \cite{},  
, paraphrasing \cite{petrovicNAACL12,ganitkevitch13Paraphrase}, 
and diversifying search results \cite{Song11SIGIR}.

In this research,  we will take advantage of large-scale query logs 
by finding nearest neighbors (queries) for a user given query from all 
queries from a big query logs. Each query point also has an associated feature vector, that is
very high dimensional and sparse $\--$ these are the set of webpages (urls) that get clicked for this query term.

@@

The motivated behind this research is to find nearest neighbors by doing a 
\emph{small} number of comparisons (sublinear in dataset size), instead of brute force linear search. 
In addition to \emph{small} number of comparisons, we also want to 
return a set of neighbor candidates with a $100\%$ precision and a large recall. 
The method we propose meet all these criterions. We do this by exploiting 
existing research in Locality Sensitive Hashing 
\cite{Indyk98STOC,Charikar02STOC,Andoni06FOCS,Andoni08CACM} and propose their 
novel variants. In particular, we develop the framework of  \lsh algorithms on a distributed 
system (e.g. Hadoop) to take advantage of its computing efficiency.

Our work includes following contributions: 
\begin{enumerate}
\item We present a distributed system (Hadoop) that can approximately find nearest neighbors for a given query from a commercial large-scale query logs in sublinear time. % Can we claim something more specific wrt time (e.g. 0(N logN))?
\item We present vanilla \lsh algorithm based on the seminal research of Andoni and Indyk  \shortcite{Andoni08CACM}. To best of our knowledge, this is the first paper that applies this algorithm to NLP applications. 
\item We propose four novel variants of vanilla LSH motivated by the research on Multi-Probe LSH \cite{LvVLDB07}.  We show that two of our variants get signiﬁcantly better recall than the vanilla \lsh by using the same number of hash tables. The main idea behind these variants is to intelligently look up multiple buckets within a table that have a high probability of containing the nearest neighbors of a query.  
%: Random Flipping and Query-Driven Flipping. 
\item  We show that the applicability of our system on two real-world applications like finding related queries (Increasing  direct display coverage is related to finding related queries) and ??.
\end{enumerate}

>>>>

 \section{Introduction}
Millions of users visit commercial search engines and ``query'' 
their interests. To provide users with high quality of services, 
search engines such as Bing, Google, and Yahoo require
intelligent analysis to realize users' implicit intents, in particular taking
advantage of large scaled query logs.  One of the key
interesting tasks involved in learning users' implicit intents often involves 
computing nearest neighbors (queries) for a user given query from all queries. 
Computing nearest neighbors is useful for 
many search-related problems on the Web and the Mobile such as 
finding related queries \cite{Jones06WWW,Jain11SIGIR,Song12WSDM}, 
finding near-duplicate queries \cite{}, spelling correction \cite{},  
, paraphrasing \cite{petrovicNAACL12,ganitkevitch13Paraphrase}, 
and diversifying search results \cite{Song11SIGIR}.

In this research,  we will take advantage of large-scale query logs 
by finding nearest neighbors (queries) for a user given query from all 
queries from a big query logs. Each query term also has an associated feature vector, that is
 high dimensional and sparse $\--$ these are the set of webpages (urls) that get clicked for this query term.

@@

The motivation behind this research is to find nearest neighbors by doing a 
\emph{small} number of comparisons (sublinear in dataset size), instead of brute force linear search. 
In addition to \emph{small} number of comparisons, we also want to 
return a set of neighbor candidates with a $100\%$ precision and a large recall. 
We can not tolerate false positives. We can afford false negatives as there 
are many relevant queries; not retrieving all of them can be sufficient for a specific problem.

The method we propose meet all these criterions. We do this by exploiting 
existing research in Locality Sensitive Hashing 
\cite{Indyk98STOC,Charikar02STOC,Andoni06FOCS,Andoni08CACM} and propose their 
novel variants. In particular, we develop the framework of  \lsh algorithms on a distributed 
system (e.g. Hadoop) to take advantage of its computing efficiency.

Our work includes following contributions: 
\begin{enumerate}
\item We present a distributed system (Hadoop) that can approximately find nearest neighbors for a given query from a commercial large-scale query logs in sublinear time. % Can we claim something more specific wrt time (e.g. 0(N logN))?
\item We present vanilla \lsh algorithm based on the seminal research of Andoni and Indyk  \shortcite{Andoni08CACM}. To best of our knowledge, this is the first paper that applies this algorithm to NLP applications. 
\item We propose four novel variants of vanilla LSH motivated by the research on Multi-Probe LSH \cite{LvVLDB07}.  We show that two of our variants achieve significantly better recall than the vanilla \lsh by using the same number of hash tables. The main idea behind these variants is to intelligently look up multiple buckets within a table that have a high probability of containing the nearest neighbors of a query.  
%: Random Flipping and Query-Driven Flipping. 
\item  We show that the applicability of our system on two real-world applications like finding related queries (Increasing  direct display coverage is related to finding related queries) and ??.
\end{enumerate}
